---
title: "Covid91 vaccine study Final2024"
author: "Hailey Watkins-Oliver!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
Males <- guy %>%
  group_by(infected,treatment) %>%
  summarize(count=n()) %>% 
  mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The population of men (males) and their association to COVID-19 by vaccination or placebo are represented in this plot. 


Plotting indicates that COVID-19 was significantly lower in those who received the immunization than in those who received a placebo. The size difference between the placebo and the vaccination is roughly 3:1.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

The percentage of boys who contracted COVID-19 after receiving the immunization was a low 26.2%. 73.8% of those who are not vaccinated contract COVID-19. The vaccine appears to have a bright future.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```


The output of the Fisher's Exact test is the data that should be closely monitored among all of these findings. The p-value, which is 2.2e-16 0.00000000000000022, is quite small. This figure indicates that it is unlikely that these outcomes happened by accident.

When compared to a placebo, our odds ratio of roughly 2.8 indicates that this vaccine is beneficial in halting COVID-19.
```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`


The odds ratio of the males with vaccines vs those with placebos has a ratio greater than 70 meaning we have a effective vaccine for males. 

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions